---
title: "Coursera Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading and preprocessing the data

```{r cars}
d <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?

```{r d, echo=TRUE}
perday <- with(d, aggregate(steps ~date, FUN=sum, na.rm=TRUE))
hist(perday$steps, main="Steps taken each day", xlab="Number of steps", ylab="Frequency")
meanPerDay <- as.integer(mean(perday$steps))
medianPerDay <- as.integer(median(perday$steps))
```


The mean is `r meanPerDay` and the median is `r medianPerDay`


## What is the average daily activity pattern?

```{r e, echo=TRUE}
perinterval <- with(d, aggregate(steps ~interval, FUN=mean))
with(perinterval, plot(interval, steps, type="l"))
intervalWithMostSteps <- perinterval$interval[perinterval$steps==max(perinterval$steps)]
```

The interval with most steps is `r intervalWithMostSteps`


## Imputing missing values

```{r f, echo=TRUE}
completeCases <- sum(complete.cases(d))
missingCases <- nrow(d) - completeCases
d1 <- d
meanperday <- with(d, aggregate(steps ~date, FUN=mean, na.rm=TRUE))
# replace missing values by the mean for that day
d1$steps <- ifelse(is.na(d1$steps), meanperday$steps[d1$date], d1$steps)
# it appears some days only have missing values so above will still leave us with some missing values which we fill with global mean
d1$steps <- ifelse(is.na(d1$steps), mean(meanperday$steps), d1$steps)

perday <- with(d1, aggregate(steps ~date, FUN=sum, na.rm=TRUE))
hist(perday$steps, main="Steps taken each day", xlab="Number of steps", ylab="Frequency")
meanPerDay2 <- as.integer(mean(perday$steps))
medianPerDay2 <- as.integer(median(perday$steps))
meanChange <- round((meanPerDay2-meanPerDay)/meanPerDay*100, digits=2)
medianChange <- round((medianPerDay2-medianPerDay)/medianPerDay*100, digits=2)
```

There are `r missingCases` missing cases

The mean is `r meanPerDay2` and the median is `r medianPerDay2`

The mean with impued missing values differs by `r meanChange`%

The median with imputed missing values differs by `r medianChange`%


## Are there differences in activity patterns between weekdays and weekends?

```{r g, echo=TRUE}
d1$date <- as.Date(d1$date)
d1$w <- factor(c("weekday","weekend"))
d1$w <- ifelse(weekdays(d1$date)=="Saturday" | weekdays(d1$date)=="Sunday", "weekend", "weekday")
d1weekday <- d1[d1$w=="weekday",]
d1weekend <- d1[d1$w=="weekend",]
perintervalWeekday <- with(d1weekday, aggregate(steps ~interval, FUN=mean))
perintervalWeekend <- with(d1weekend, aggregate(steps ~interval, FUN=mean))
par(mfrow=c(1,2))
with(perintervalWeekday, plot(interval, steps, type="l", main="Activity on weekdays"))
with(perintervalWeekend, plot(interval, steps, type="l", main="Activity on weekends"))
```